Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-18 doc/GMP2015/s03_drp.tex
Further proof reading. Fixed \emph problem!
Expanded on the DR-planning example and algorithm. Minor change to recomb notation.

However, to guarantee that any two are \vemph{solutions} to $OptimalDRP(G)$, it must ensure minimum fan-in for any descendant. Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$. Suppose we select $i$ and $j$, where $i\neq j$, as the children. Since
% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
% the children of this node will be
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
% and
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
$C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
the children of this node will be
$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
and
$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?} This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,I\cup R_k)$ for $k\in S_N$, with every $k$ appearing at least once. Thus, regardless of the choice of $i,j$ and then $m,n$ etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
\end{enumerate}

This proof then applies to itself recursively to show that the fan-in of the children will also be minimum, thereby making this the optimal DR-plan.
\end{proof}

@@

For (4), because subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained, all cases have already been exhausted.
% \begin{enumerate}
%     \item If $F_i\cup F_j$ were trivial, then, by definition, $F_i$ and $F_j$ must be trivial.

@@

%     \item Subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained. All cases are already exhausted.
% \end{enumerate}
\end{proof}

@@

\begin{proof}
%I wouldn't do the forward and backward direction italics, or say proof by contradiction. Just start the first paragraph with Suppose C_i union C_j is well constained maximal proper and assume it is not C.
Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and $C_i,C_j$ were not vertex-maximal to begin with.

In the reverse direction, we know $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem). Thus, $C_i\cup C_j=C$ is well-constrained.
\end{proof}

@@

\subsection{Examples}

@@

\caption{(\ref{fig:c2c3ofk33s:a}) A doublet with each edge of the triangles replaced by a $K_{3,3}$. This pattern continues for a total of $N$ triangles. (\ref{fig:c2c3ofk33s:b}) The DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes. $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs. The dashed lines indicated that this exact structure is repeated until only $\{a_{(N+1)/2},b_{(N+1)/2},c_{(N+1)/2}\}$ remains in the case that $N$ is odd, or until only $\{a_{N/2},b_{N/2},c_{N/2},a_{(N/2)+1},b_{(N/2)+1},c_{(N/2)+1}\}$ remains if $N$ is even.}
\label{fig:c2c3ofk33s}
\end{figure*}

@@

\todo{Update, explain pebble game and frontier, complexity}

The algorithm to find the optimal DR-plan is a straightforward application of the theory. It begins by finding the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph. If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph. Otherwise, all the subgraphs are the children. Apply recursively to the children. This naturally terminates when you reach trivial subgraphs.

>>>>

However, to guarantee that any two are \vemph{solutions} to $OptimalDRP(G)$, it must ensure minimum fan-in for any descendant. Take the set $S_N=\{1,\dots,N\}$, then we call $I=\bigcap_{k\in S_N}{C_k}$ and $R_k=C\setminus C_k$. Suppose we select $i$ and $j$, where $i\neq j$, as the children. Since
% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
% the children of this node will be
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
% and
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
$C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
the children of this node will be
$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
and
$Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?} This continues for $N-1$ levels total, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $Idc(C,I\cup R_k)$ for $k\in S_N$, with every $k$ appearing at least once. Thus, regardless of the choice of $i,j$ and then $m,n$ etc., the DR-plan has fan-in of two for every node for the next $N$ levels, at which point the nodes contain the same subgraphs.
\end{enumerate}
%
This proof then applies to itself recursively to show that the fan-in of the children will also be minimum, thereby making this the optimal DR-plan.
\end{proof}

@@

For (4), because subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained, all cases have already been exhausted.
\end{proof}

% \begin{proof}
% Use the fact that, by definition, $d(F_i)=k$ and $d(F_j)=k$. Also, the equation $d(F_i\cap F_j)=d(F_i)+d(F_j)-d(F_i\cup F_j)$.
% \begin{enumerate}
%     \item If $F_i\cup F_j$ were trivial, then, by definition, $F_i$ and $F_j$ must be trivial.

@@

%     \item Subgraphs of a well-constrained graph can only be trivial, under-, or well-constrained. All cases are already exhausted.
% \end{enumerate}
% \end{proof}

@@

\begin{proof}
Assume $C_i\cup C_j \neq C$, then we just found a larger proper subgraph and $C_i,C_j$ were not vertex-maximal to begin with.
%
In the reverse direction, we know $C$ is either some non-leaf node (well-constrained by definition of a DR-plan) or $G$ itself (well-constrained by definition of the problem). Thus, $C_i\cup C_j=C$ is well-constrained.
\end{proof}

@@

\subsection{Example}

@@

\caption{(\ref{fig:c2c3ofk33s:a}) A doublet ($C_2 \times C_3$) with each edge of the triangles replaced by a $K_{3,3}$. This pattern continues inwards for a total of $N$ triangles, indicated by the dashed lines. (\ref{fig:c2c3ofk33s:b}) Most of the DR-plan of this graph, omitting further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges and of edges into the component nodes. $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs. The dashed lines indicated that this exact structure is repeated.}
\label{fig:c2c3ofk33s}
\end{figure*}

This section details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}. The graph $G$ has only two isostatic vertex-maximal subgraphs, $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle $N$. These intersect on $G$ without triangle $1$ and $N$ which is clearly well-constrained. As explained in the proof of Theorem \ref{theorem:main}, part 2, since there are only 2 possible children, their intersection must occur 2 levels below the parent. Just as expected, it is on the third level, as it is a child of both of $G$'s children. Furthermore, it intersects trivially with the edges that connect triangle $1$ and $N$ to $2$ and $N-1$, respectively; these necessarily intersect with triangle $1$ and $N$ themselves trivially (by Theorem \ref{theorem:main}, part 1) making all of these children. The triangles then decompose into their constituent $K_{3,3}$ graphs, which then decompose into 9 edges.

Now, the self-similar nature of the graph is clear; $G$ without triangle $1$ and $N$ has the same structure, and thus the same pattern is repeated until only triangle $(N+1)/2$ remains in the case that $N$ is odd, or until only triangles $N/2$ and $N/2+1$ remain if $N$ is even.

@@

The first step of the algorithm to find an optimal DR-plan is the most difficult --- to find the well-constrained vertex-maximal proper subgraphs of the input well-constrained graph. There are two approaches: (1) the bottom-up approach, using the Frontier algorithm \uncited to grow these components from single vertices; and (2) the top-down approach, using the pebble game algorithm \uncited on each node and taking the largest components that are not subgraphs of any larger ones. After the subgraphs have been found, it is a straightforward application of the theory. If the intersection of any two of the subgraphs is well-constrained, then take those two to be children of the graph. Otherwise, all the subgraphs are the children. Apply recursively to the children. This naturally terminates when you reach trivial subgraphs. The complexity of this depends on the algorithm for finding children (both are polynomial) and the depth of the plan (the deeper the plan, i.e.\ the smaller the max fan-in, the longer the running time). \todo{Think more about complexity.}

### 2014-12-18 doc/GMP2015/s04_recombination.tex
Further proof reading. Fixed \emph problem!
Expanded on the DR-planning example and algorithm. Minor change to recomb notation.

\begin{proof}
The proof follows from the fact that 2-trees are well decomposable and
have simple DR-plans of size 2. We know that $G$ can be reduced by
removing $k$ edges to create a partial-2-tree $G'$ which can then be
completed to an (isostatic) 2-tree by adding some set of non-edges
$F$. Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2,
proving the theorem.
\end{proof}.

@@

\begin{proof}
    The Cartesian realization space of $(H,<\delta_E', \lambda_F>)$ is
    computed easily with a DR-plan of size 2, and is the union of
    $2^t$ solutions (modulo orientation preserving isometries) each
    with a distinct orientation type, where $t$ is the number of
    triangles in the 2-tree $H$; here $\delta_{S}$ is the restriction
    of the length vector $\delta$ to the edges in $S$. A desired
    solution $p$ (or connected component of a solution space) of
    $(G,\delta)$ of an orientation type $\sigma_p$ can be found by a
    subdivided binary search   of the cartesian realization space of
    $(H, <\delta_{E'},\lambda_F>)$ of orientation type $\sigma_p$ - as
    $\lambda_F$ ranges over the discretized convex polytope
    $\Phi_F(G',\delta_E')$ with bounding hyperplanes described in
    Theorem \ref{theorem:convexcayley}. A solution $p$  is found  when
    the lengths for nonedges in $D$ match $\delta_D$.
\end{proof}

@@

\medskip\noindent
\textbf{Example}. \todo{This uses different example format, not their environment. Should this, or section 3 change?}
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
isostatic graphs, i.e, its DR-plan has a root and 9 children
corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$ corresponding to the root of this DR-plan involves
solving a quadratic system with 8 equations and variables. Instead of
simultaneously solving this system, we could instead use the fact that
$G=K_{3,3}$ is in OMD$_1$: remove the edge $e$ in Figure
\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
interval of attainable length values $\lambda_F$ for the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or
coincidences in the realizations of $(G,\delta)$, the realization
space of $(H, <\delta_{E\setminus e}, \lambda_f>)$ has 16 solutions
$q^p_{\lambda_f}$ (modulo orientation preserving isometries), with
distinct orientation types $\sigma_p$ (two orientation choices for
each of the 4 triangles) that can be obtained by solving a sequence of
4 single quadratics in 1 variable (DR-plan of size 2). By subdivided
binary search in the interval $\lambda_f \in \Phi_f(G',
\delta_{E\setminus e})$, the desired solution $p$  of $(G,\delta)$ is
found when the length of the nonedge $e$ in  the realization
$q^p_{\lambda_f}$ is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{sec:tdecomp}
%
It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
the number of edges to be removed to ensure an isostatic completion
that is decomposable with a small DR-plan) by considering reduced
graphs $G'$ (and modified graphs $H$) that come from a larger class
than partial 2-trees but nevertheless have convex Cayley configuration
spaces at least when the realization space is restricted to a
sufficiently comprehensive orientation type. In particular, the so-
called \dfn{tree-decomposable graphs of low Cayley complexity}
\uncited include the partial 2-trees and many others that are not
partial 2-trees. These too result in DR-plans of size 2 or 3, putting
$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
Criterion (c) is met---for example when $k=1$---because 1-dof Cayley
configuration spaces of linkages based on such graphs $G'$ are known
to be single intervals when a comprehensive orientation type
$\sigma_p$ of the sought solution $p$ is given. In addition, the
bounds of these intervals are of low algebraic complexity. More
precisely, the bounds  can themselves be computed using a DR-plan of
size 2 or 3, i.e, the computation of these bounds is tree-
decomposable. Alternatively, the bounds are in a simple quadratic or
radically solvable extension field over the rationals, or can be
computed by solving a triangularized system of quadratics.

>>>>

\begin{proof}
The proof follows from the fact that 2-trees are well decomposable and
have simple DR-plans of size 2. We know that $G$ can be reduced by
removing $k$ edges to create a partial-2-tree $G'$ which can then be
completed to an (isostatic) 2-tree by adding some set of non-edges
$F$. Thus the modified graph $H = G'\cup F$ has  a DR-plan of size 2,
proving the theorem.
\end{proof}

@@

\begin{proof}
    The Cartesian realization space of $(H,\left<\delta_{E'},
    \lambda_F\right>)$ is computed easily with a DR-plan of size 2,
    and is the union of $2^t$ solutions (modulo orientation preserving
    isometries) each with a distinct orientation type, where $t$ is
    the number of triangles in the 2-tree $H$; here $\delta_{S}$ is
    the restriction of the length vector $\delta$ to the edges in $S$.
    A desired solution $p$ (or connected component of a solution
    space) of $(G,\delta)$ of an orientation type $\sigma_p$ can be
    found by a subdivided binary search   of the cartesian realization
    space of $(H, \left<\delta_{E'}, \lambda_F\right>)$ of orientation
    type $\sigma_p$ - as $\lambda_F$ ranges over the discretized
    convex polytope $\Phi_F(G',\delta_E')$ with bounding hyperplanes
    described in Theorem \ref{theorem:convexcayley}. A solution $p$
    is found  when the lengths for nonedges in $D$ match $\delta_D$.
\end{proof}

@@

\medskip\noindent
\textbf{Example}. \todo{This uses different example format, not their environment. Should this, or section 3 change?}
A graph $G=K_{3,3}$  cannot be decomposed into any nontrivial
isostatic graphs, i.e, its DR-plan has a root and 9 children
corresponding to the 9 edges. Solving or recombining the system
$(G,\delta)$ corresponding to the root of this DR-plan involves
solving a quadratic system with 8 equations and variables. Instead of
simultaneously solving this system, we could instead use the fact that
$G=K_{3,3}$ is in OMD$_1$: remove the edge $e$ in Figure
\ref{fig:omd_k33_example} to give a partial 2-tree $G'$. Now add the
nonedge $f$ to give a 2-tree $H$ with a DR-plan of size 2. The Cayley
configuration space $\Phi_f(G', \delta_{E\setminus e})$ is a single
interval of attainable length values $\lambda_F$ for the edge $f$.
When $\delta$ is generic, i.e, does not admit collinearities or
coincidences in the realizations of $(G,\delta)$, the realization
space of $(H, \left<\delta_{E\setminus e}, \lambda_f\right>)$ has 16
solutions $q^p_{\lambda_f}$ (modulo orientation preserving
isometries), with distinct orientation types $\sigma_p$ (two
orientation choices for each of the 4 triangles) that can be obtained
by solving a sequence of 4 single quadratics in 1 variable (DR-plan of
size 2). By subdivided binary search in the interval $\lambda_f \in
\Phi_f(G', \delta_{E\setminus e})$, the desired solution $p$  of
$(G,\delta)$ is found when the length of the nonedge $e$ in  the
realization $q^p_{\lambda_f}$ is $\delta_e$.
%
\subsection{Optimized Modification by Enlarging the class of Reduced
Graphs}
\label{sec:tdecomp}
%
It is possible in principle to decrease $k$ for a OMD$_k$ graph (i.e,
the number of edges to be removed to ensure an isostatic completion
that is decomposable with a small DR-plan) by considering reduced
graphs $G'$ (and modified graphs $H$) that come from a larger class
than partial 2-trees but nevertheless have convex Cayley configuration
spaces at least when the realization space is restricted to a
sufficiently comprehensive orientation type. In particular, the so-
called \dfn{tree-decomposable graphs of low Cayley complexity}
\uncited include the partial 2-trees and many others that are not
partial 2-trees. These too result in DR-plans of size 2 or 3, putting
$G$ in the class OMD$_k$ and thus meeting Criteria (a) and (b). The
Criterion (c) is met---for example when $k=1$---because 1-dof Cayley
configuration spaces of linkages based on such graphs $G'$ are known
to be single intervals when a comprehensive orientation type
$\sigma_p$ of the sought solution $p$ is given. In addition, the
bounds of these intervals are of low algebraic complexity. More
precisely, the bounds  can themselves be computed using a DR-plan of
size 2 or 3, i.e, the computation of these bounds is tree-
decomposable. Alternatively, the bounds are in a simple quadratic or
radically solvable extension field over the rationals, or can be
computed by solving a triangularized system of quadratics.
